%!TEX root = ../thesis.tex
%*******************************************************************************
%****************************** Second Chapter *********************************
%*******************************************************************************

\chapter{Background}

introduction to 4P, 

Traits that are related to psychological disorders
psychological disorder symptoms

\section{Factors contributing to mental health}
% Uncomment this line, when you have siunitx package loaded.
%The SI Units for dynamic viscosity is \si{\newton\second\per\metre\squared}.
The 4P model is a biopsychosocial approach to analyze mental health development and progression \cite{}. The 4P model suggests that there are four types of factors contribute to development of all psychological disorders, each type of factors can be further divided into biological factors, psychological factors and social factors. The four types of factors are: predisposing risk factors, precipitating risk factors, perpetuating risk factors and protective factors. 

Predisposing risk factors make an individual more susceptible to a psychological disorder. For example, genetic component, gender differences, personality traits and parental modeling. Precipitating risk factors contribute to the onset of psychological disorders, for instance, traumatic event or stressful life events. Perpetuating factors maintain the symptoms span from psychological disorders, such as chronic stress, cognitive bias, social stigma, social roles. Finally, protective factors build up the resilience of an individual \cite{}. Figure \ref{tab:4P} demonstrates the relationships of these factors. Note that some of the precipitating factors can be perpetuating factors if they exist for a long time. 

We attempt to study some of the factors via the lenses of social media platform. Social media behaviors reflect psychological and social factors to some extend but it's nearly impossible to inspect biological factors from social media data, except for gender differences. Another factor that is difficult to inspect in social media at this stage is childhood experience. As a result, my PhD project will only focus on inspecting factors that are available in social media data.

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}


% section section_name (end)
\begin{table}
\caption{Biopsychosocial Approaches in 4P Model}
\label{tab:4P}
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|p{2cm}|p{4cm}|p{4cm}|}
\hline
\textbf{4P factor model}            & \multicolumn{3}{l|}{\textbf{Biopsychosocial approach}}      \\ \hline
                           & \textbf{biological factors} & \textbf{psychological factors}          & \textbf{social factors}                                      \\ \hline
\textbf{Predisposing risk factors}  & genetic, gender    & personality traits, mood stability, self-disclosure, self-esteem, perfectionism, sensitivity to stress       & parental modelling,                                 \\ \hline
\textbf{Precipitating risk factors} & physical injuries  & stress/trauma                                                                                                & change in family dynamics                           \\ \hline
\textbf{Perpetuating risk factors}  & chronic diseases   & chronic stress, malnutrition, crowding, pollution                                                            & social stigma, social roles, culture, social status \\ \hline
\textbf{Protective factors}         & genetic            & effective coping skills, high resilience, high cognition skills, high self-esteem, low sensitivity to stress & social support                                      \\ \hline
\end{tabular}%
}
\end{table}

\subsection{predisposing factors}
When individuals are exposed to a stressor, predisposing factors influence how people experience stress and how they cope with it. These factors affect an individual in the probability to encounter stressors. Note that people select their environment and shape them. They also affect people's tendency to react to situation and their tendency in coping \cite{suls2005daily,vollrath2001personality,watson1999personality}.

\subsubsection{personality}

Consistent evidence has been found that some dimensions in personality co-occurred with certain mental disorders, but there's no direct link between personality and mental health problems. Personality is largely shaped by child-rearing practices and family interactions (cite). As a result, the link between personality and mental health can be traced back to family rearing style.

There are many types of personality models, such as, Myers–Briggs Type Indicator (MBTI),Cattell’s 16 Factor Model, Big Five Model, among which the five factor personality model (Big5) is most widely adopted in personality studies. Big5 suggests five board dimensions to describe human personality:  openness to experience, conscientiousness, extraversion, agreeableness, and neuroticism. There has been consistent evidence showing that personality plays an important role in the experience of stress. Individuals with high neuroticism often experience more negative emotions (cite), and negative emotions affect how an individual cope with stress \cite{kiecolt2002emotions}. Neurotism has been found to be an increase risk factor for depression and anxiety. \cite{talley1986association,cattell1961meaning} Internet addiction often occurs in introverted people \cite{xiuqin2010mental}. However, personality and mental health do not have a direct causation relationship. There are many mediating factors between personality types and mental disorders. For example, rumination and worrying mediate neuroticism and depression \cite{roelofs2008rumination};

\subsubsection{self-disclosure}
Self-disclosure is the amount of information an individual disclose to significant others. There are three major parameters of self-disclosure: the amount of information disclosed; depth of the information disclosed and the time spent on describing the information. The evaluation of self-disclosure must involve the three parameters \cite{cozby1973self}.


Self-disclosure is important to mental health problem because it affects one's help seeking behaviors. Individuals with high self-disclosure level are more likely to seek professional help \cite{hinson1993willingness}. Self-disclosure is also a critical problem for psychotherapist. On one hand, patients often go to psychotherapy to seek solution to loneliness, which is a result of difficulty in building intimate relationships. One the other hand, patients wish to conceal themselves \cite{fisher1990shared, stricker1990self}. 

Self-disclosure is, again, shaped by the family rearing style. Fisher(1982) proposes that when infant/child involves in a communication with significant others, they receive positive or negative feedback . The feedback provided form a feedback loop. The positive feedback loop encourage a child to disclose more, whereas, the negative feedback loop does the opposite. Failing to disclose results in an accumulation of secrets, the greater the degree of secrets, the more alienation of self and others. cite() Besides family rearing style, self-disclosure is also affected by gender, culture, race and the relationship with the significant others.

\subsubsection{self-esteem}
Self-esteem is the beliefs and evaluation individuals hold about themselves \cite{burns1982self}. These beliefs are inner influences that guild and government individuals through out their lives. The development of self-esteem is dependent on support and approval from significant others and self-perceived competence in the domains that are important to self. The distance between ideal self and real self is critical to the development of self-esteem cite. 

Self-esteem is a protective factor in mental health because self-esteem and the ability to face challenge determine what happens to an individual in challenging situations. Conversely. poor self-esteem plays a critical role in the development of mental disorders. \cite{mann2004self}. High self-esteem is related to better life satisfaction \cite{zimmerman2000self}.  Self-esteem is a strong predictor of subjective well-being \cite{furnham2000perceived}. 

Whereas, Low self-esteem lead to maladjustment \cite{garmezy1984study}, depressed mood and depressive disorders \cite{rice1998self,dori1999depression}. Self-esteem alone has no direct contribution to depression, but self-esteem interact with accumulate stress highly predicts depression \cite{miller1989self}.  Shin (Shin, 1993) found that when self-esteem, social support and cumulative stress are introduced to the regression analysis of depression, only self-esteem accounts for significant additional variance \cite{shin1993factors}. 

\subsubsection{gender and cultural differences}

A large amount of literature has found significant gender differences in depressive and anxiety disorders. Nevertheless, gender is an influential factor to help-seeking behavior. One one hand, women in genderal have higher self-disclosure level, thus they are more willing to seek help. On the other hand, gender stereotype such as women are prone to affective disorder and men are prone to substance use disorder also reinforce the stigma and affect the help seeking behaviors. Gender differences also exist in the comorbid rate of depression and anxiety \cite{afifi2007gender}. 

Cultural differences also influence the development of mental disorders. There are cultural bound syndromes, which are disorders that are more likely to occur in certain culture and context. For example, obesity and anorexia nervosa in the United States \cite{ritenbaugh1982obesity}, social withdrawal in Japan \cite{teo2010hikikomori}. Cultural stigma also restraint the help-seeking behaviours of people from a certain culture. A person live in a culture full of stigmatizing image of mental health may internalize these ideas and suffer from a diminish in self-esteem and confidence \cite{corrigan1998impact,holmes1998individual}. 

In summary, a lot of the predisposing factors, such as personality, self-disclosure and self-esteem, are dependent on family rearing style. It's nearly impossible to extract childhood information of a person from social media data, because social media only exploded in the recent decade. However, it's possible to extract information that reflects predisposing risk factors from social media data. As a result, my project will involve studying some of these disposing factors. Predisposing factors makes up the vulnerabilty of mental disorders, but predisposing factors alone are not directly link to mental disorders. Their interaction with precipitating factors - stress, are most predictive of mental health.


\subsection{precipitating factors and perpetuating factors}

Life stress has been acknowledged as an important factors in illness onset (citation). It has been found that major life events, such as divorce, death of a closed friend or relative, being fired from work and so on are precipitating factors on mental illnesses (citation). However, the correlations between stress and health outcome is confound by many factors, such as the predisposing factors in the previous sections. As a result, the correlations between stress and health are modest (r =  0.2 - 0.4)  The correlations are much stronger among those who are already ill \cite{aldwin2004interface}.

\subsubsection{measurement of stress}
Most of the literature analyze stress from two perspective: psychological and environmental perspectives. The psychological perspective focuses on the individual's appraisal of their ability to cope with stress, while the environmental perspective focuses on the life events that cause the stress \cite{mcgrath1970social}. 

However, measuring stress level with the type of stressor is far from accurate due to individual differences in assigning meanings to life events. Objective rating of life events, for instance, as{}suming that bereavement must be extremely stressful, does not account for the individual differences.  Wortman and Silver found that one-third of the bereaved spouses do not appear to be distressed at any point up to one year after the spouse's death \cite{wortman1989myths}. Thus, subjective ratings of stress are more predictive to mental health conditions.  Scales that measure the subjective rating of stress include Schedule of Recent Life Experiences (SRE) \cite{nuckolls1972psychosocial}, the Social Readjustment Rating Scale (SRRS) and so on (cite). Another way to measure stress is the interview. Interview measurement is particularly critical in diagnosis because the timing of exposure to the stressor indicate the onset of the illness.


\subsubsection{types of stress}
\textbf{Random stressors and systematic stressors}. Random stressors occur with relatively equal probability across various social group, e.g. miscarriages. Systematic stressors happen in specific social location or social group, e.g being robbed. 

\textbf{Chronic and acute stressors.}  Acute stressors result in onset of illness, while chronic stressors expose an individual to long-term stress. Both of them play an important role in the development of mental disorders. However, sometimes the boundary of chronic and accute stressors are not clear cut. In fact, many negative events are not uniformily evaluated as discrete or chronic \cite{avison1988stressful}. For instance, a car accident is an acute stressor, but the injuries follow with the accident can take years to recover, which makes this accident a chronic stressor.

\textbf{Change events and negative change events.} There are consistent findings that supports a positive relationship between negative change event and mental illness. A large amount of studies shows that negative events rather than positive events precipitate symptoms \cite{rabkin1976life,turner1995epidemiology}. Whereas, the findings for positive event and mental illness are contradictory \cite{zautra1983life}. As a result, life event checklists emphasis the undesirability as the most important factor to mental illness. 

\subsubsection{Other aspects of stress} 
\textbf{Uncontrollability of event.} It is also important to note that uncontrollable event is more stressful than those that are under an individual's control. For example, initiate a breakup with a partner is less stressful than being told the bad news by a partner; Being fired is more stressful than quitting a job. 


\textbf{culmulative of stress and role strain.} It is believed that the accumulation of stress is more stressful than an isolated life events \cite{evans2013childhood}. The literature suggests that the duration of stress effect usually lasts for 6 months  \cite{depue1986conceptualization,norris1987transitory}. In addition, chronic role strain and hassles may also increase the severity of the illness. 
 
Chronic role strain is a systematic stressor, which emphasis the social causes of psychological distress. Intuitively, social role often brings chronic role strain, for example, the role of a working mother need to thrive in her career and take care of children at home. As a consequence, in the measurement of stress, respondent who take on more social roles should report higher tension in stressful events. However, social role is, in fact, a confunding facter, because the number of roles one taken is linked to one's social competence. Social competence is negatively correlated with distress. 


\subsection{protective factors}
\subsubsection{life experience}
The earliest literature of psychology has discovered the relationship with life expereince and psychological disorders. Later on, a considerable body of reseach find that life experience varied in their risks to develop mental disorders. Even for children who have experience the most severe  stressors, only half of them succumb to the adversities \cite{rutter1979protective}. The risks of life experience is mediated by predisposing factors and protective factors. 

Adverse life experience itself can be a protective factor. There is large amount of animal studies show that physical stresses in early life lead to neurological changes that improve the animal's resistence to stress happen later in life \cite{hennessy1979stress}. In certain circumstances, adversity enhance an individual's psychological hardness. Whether one will develop psychological hardness depends on predisposing factors and timing of the event. Timing and meaning are important as timing affects the meaning attached to the event and meaning affects individuals appraisal of whether the event is postive or threatening \cite{rutter1985resilience}

\subsubsection{cognition}
People's cognition affect their appraisal of their ability to cope with the adverse situation. On one hand, cognitive skills are linked to the development of self-esteem. Whether one is successful in the domain of importance is crucial to the development of self-esteem and this is largely dependent on cognitive skills. On the other hand, coping stragies differentiate people who think they can cope with the situation and those who can't. The characteristics of many people who have experience chronic stress is that they feel helpless and they can't do anything to change the situation \cite{rutter1985resilience}. 


\subsubsection{social support}

Social embeddedness, perceived support and enacted support are three major aspects of social support. Social embeddedness refers to an individual's connections to significant others and may include one's role in the family or the community and one's social network. Social embeddedness both create stress and provide support. On one hand, social roles can lead to role strains but reflect competence, which is an important component in self-esteem. On the other hand, tensions in social ties create stress, but healthy social ties brings social support. As a consequence, social embedness, which is reflected by the social network on social media could be a confounding factor in studying social support. 

Enacted social support refers to the support that other people provide. Perceived social support refers to whether an individual perceive that they have received the support, it focuses on the adequacy and availability of the support.  Perceived support is more likely to have stress-buffering effect because they influence the appraisal of stressful situations \cite{cohen1984social}, as well as by Cutrona and her colleagues \cite{cutrona1990type}. As a result, whether one perceive they receive support is more important to health than whether they actually receive support. 


\textbf{Types of social support.} The types of social support are slightly different in various theories. Gottlieb identified four classes of social support from 26  26 categories of helping behaviors \cite{gottlieb1978development}. The four classes are emotionally sustaining behaviors, problem-solving behaviors, indirect personal influence, and environmental action. House and colleagues proposed three major types of support: emotional, informational, and tangible support \cite{house1985social}. Cutrona and colleagues proposed esteem, network (companionship), informational, tangible, and emotional support \cite{cutrona1990interpersonal}.


\subsection{Conclusion}
Summary, protective factors may be positive or negative life experience, or may not be an experience at all. It could be the quality of a person, such as the predisposing factors (self-esteem, personality, self-disclosure) or it could be social support. All these factors interact with stress in predicting the risk of mental illness. Hence, the study of stress's effect on mental health can not be isolated from the factors that entangles with it. 


\section{Psychological disorders and their symptoms}

The symptoms in psychological disorders are considered as presenting factor in addition to the 4P model, which expand the 4P model to a 5P model. The presenting factor refers to the obvious signs or symptoms in the assessment.

\subsection{Diagnostic criteria}
The aim of my research is to study symptom related behaviours on social media but not for diagnosis. Hence, identifying symptoms from the diagnostic manual is one of the crucial steps. The Diagnostic and Statistical Manual of Mental Disorders, Fifth Edition (DSM-5) is the latest version of diagnostic criteria manual \cite{american2013diagnostic}. DSM-5 diagnostic criteria contain 22 categories of disorders, some of the disorders has been widely studied in the social media context. There are computer classifiers that claim to be able to diagnose disorders automatically.

However, the diagnosis of mental disorders is extremely complicated. First of all, there are multiple subcategories under each major disorder categories. For example, depressive disorders include disruptive mood dysregulation disorder, major depressive disorder, persistent depressive disorder, premenstrual dysphoric disorder and so on. Diagnostic criteria between the subcategories are very similar, the distinction between the depressive disorder subcategories mainly lies in the onset time and how long the symptoms prolong. 

Second, a major concern of diagnosis is co-morbidity, which means an individual is diagnosed with more than one disorders at the same time. Kessler and others conduct a National Co-morbidity Survey and found that the 'pure' cases are very rare \cite{kessler1994lifetime}. Schwalberg et al. (1992) found the comorbidity between eating disorders and anxiety is 70–80\% \cite{schwalberg1992comparison}, Judd and colleagues found the co-morbidity between general anxiety disorder and depression is 80\% \cite{judd1998comorbidity}. The co-morbidity between bipolar disorder and another psychological disorder is 61\%. \cite{raja2004clinical}. The comorbidity may refelct a common vulnerability. For instance, poor coping skills  \cite{andrews1996comorbidity} is one of the vulnerabilities that expose an individual to higher risks of anxiety and depressive disorders. 

Nevertheless, DSM-5 also stress culture and gender related diagnostic issues. In many cultures, somatic symptoms are predominant symptoms in depressive disorders. A majority of depressive disorder cases are unrecognized in primary care settings because various culture or gender related somatic symptoms are more predominant than affective symptoms. In addition, the female-to-male prevalence ratio, ages of onset various across cultures \cite{harvey2004cognitive}. 

\subsection{Symptoms}

Social media behaviours reflect some of the symptomatic behaviours. In order to study psychological disorders via social media data, we need to understand symptoms that often occur in different disorders. We analysis symptoms in some of the psychological disorders that are often studied with social media data \cite{american2013diagnostic}.

\textbf{Depressive disorders.} Some of the major symptoms included in the diagnostic criteria and occur in most of the subcatergories include: depressed mood, loss of interest or pleasure, insomnia or hypersonmnia, fatigue, anxiety, irritability and anger \cite{american2013diagnostic}.


\textbf{Anxiety disorders.} The distinctions between different subcategories under anxiety disorders lies in the themes or context that trigger the symptoms. The symptoms include,  excessive fear or anxiety concerning the theme, repeated nightmares involving the theme, repeated complaint of physical symptoms (e.g. headache, stomachache). A common feature of anxiety disorders is that the fear, anxiety or avoidence is persistent for a long time \cite{american2013diagnostic}.


\textbf{Bipolar disorder.} Bipolar disorders include a manic epispde and a major depressive episode. The distinctions between different subcategories in bipolar disorder lies in the frequency of the manic-depressive cycles. During the manic episode, patient has an inflated self-esteem, thought racing, distractibility and decreased need for sleep. The major depressive episode is similar to depressive disorders. Mood swing is a common symptom exsits in all stages of bipolar disorder \cite{american2013diagnostic}.

\textbf{Stress and Trauma related disorder.} Minimal social and emotional responsiveness to others, limited positive affect. Irritability, sadness or fearful of stimuli related to the stressful event. Intrusion symptoms such as receurrent memories or distressing dreams of of the traumatic events. Symptoms that are specific related to post-traumatic stress disorder (PTSD) include distorted cognitions about the cause and consequences of the traumatic event and persistent negative emotional state. Loss of interests, irritability, hypervigilance, sleep disturbance, problem with concentration and so on \cite{american2013diagnostic}.

\subsection{symptoms and social media behavior}

Different social media platforms has various types of functionalities. Common functionalities include posts, comment, likes of the posts. We will discuss the differences between social media platform in the next section. Some of the symptoms mentioned in the previous section are often studied in social media platform, for instance, the affect of the post, time stamp of the posts as an indicator of sleep pattern, anger and other emotions indicated in the text. Since mood changes is one of the prominant features that we can identify with social media text, hence psychological disorders feature with affective symptoms are well studied with social media data, such as schizophrenia, bipolar, depression, and anxiety disorders cite(). The Linguistic Inquiry and Word Count (LIWC) is often used in the computer text analysis. It consists of more than 100 topics that indicate the psychological processing to the text author. 
 
Other social media functionalities related to symptoms or psychological states including Likes, number of friends, number of comments to the posts. XX et al found that likes of posts on Facebook and Twitter reflects insterests and psychological states of a user (cite). Number of friends as an index for social embeddedness cite(). However, using social network as a feature to indicate social embeddeness and social support could have a mixed effect, as we mentioned in the social support sections that social ties can bring stress or provide social support. Number of comments of the posts is used as an indicator of social support in studies on reddit. However, number of comments only relfects some of the comments or the post itself is controversial or debatable, we will discuss this issue in Chapter 3. In addition, distorted cognition sometimes can be identified in language, but it hasn't been studied with social media data. We will explore distorted cognition in Chapter 4. 

Symptoms can be observed in social media data include affect and sleep patterns. Occassionally social media users mention drugs they are taking or specific physical symptoms they experienced on social media platforms. However, the disclosure of this kind of information is rather rare in social media platform such as Facebook and Twitter, it's more often on anonimous platform such as Reddit. We will explore the differences between social media platform in the social media study on mental health section.


\subsection{transdisgnostics framework and RDoC}

Biological theories of psychological disorders find some disorders have similar b,rain s,tructures and neurotransmitters. For example, amydala is related to depression \cite{bowley2002low}, schizophrenia \cite{kosaka2002differential,kubicki2002voxel} , obsessive-compulsive disorder \cite{szeszko1999orbital}, and post-traumatic stress disorder \cite{rauch2000exaggerated}. Whereas, the diagnostic criteria cannot explain the high ratio of comorbidity. It is a categorical system that gives a clear cut-off line to disorders based on the observation of symptoms. However, the symptoms of various disorders often overlap, which result in arbitrarily defined boundaries of the disorders. 
,
Recently, the emergence of transdiagnostic perspective understand psychological disorders outside of the traditional framework. It focuses on identifying the common factors that maintain the behaviors or symtpoms of psychological disorders. The underlying constructs include attention, memory, reasoning, thought and behaviors \cite{harvey2004cognitive}. A transdiagnostic approaches to treatment is therapies that target a wide range of diagnosis without tailoring the protocol to specific diagnosis \cite{mansell2008cognitive}. A large amount of studies found the transdiagnotic approach is no less effective than tranditional CBT cite().

Corresponding with the transdiagnostic treatment and research, in 2008, the US National Institute of Mental Health initiate the The Research Domain Criteria (RDoC) project in contrast to the  Diagnostic and Statistical Manual of Mental Disorders in an attempt to link the current diagnostic criteria to the underlyting neurobiological systems. The comorbidity of mental disorders might reflect different patterns of symptoms have shared risk factors and underlying disease factores. cite(Research Domain Criteria (RDoC): Toward a New Classification Framework for Research on Mental Disorders). The RDoc defines six constructs: Negative Valence Systems (Fear, Anxiety, Loss, Frustrative Nonreward), Positive Valence Systems (Reward learning, Reward valuation, Habits), Cognitive Systems (Attention, Perception, Declarative Memory, Working Memory, Cognitive control), Systems for Social Processes (Attachment formation, Social Communication, Perception of self, Perception of others), Arousal/Modulatory Systems (
Arousal, Circadian rhythm, Sleep and wakefulness) cite(https://www.nimh.nih.gov/research-priorities/rdoc/constructs/rdoc-matrix.shtml). 


\subsection{Conclusion}
Based on the complexity of diagnostic criteria and the high ratio of comorbidity, it is unlikely to build a machine learning classifer that is able to predict a task that is already very difficult for human decision. However, we can identify the RDoC constructs or the transdiagnostic components from social media data. Social media posts is a good sources for valence analysis. Social media platform such as Facebook or Twitter provide longitudinal valence data. LIWC psychological processing categories has included some of the subcategories of valence, such as anxiety and reward. The social media posts also contain information about social communication, perception of self and others. Time stamp of the social media posts might indicate the modulatory Systems if the person is a regular poster on social media platform. Meanwhile, social media posts provide a good sources for analysing thoughts in the transdiagnostics framework. As a result, our studies will not focus on using social media to infer the diagnostic criteria. Instead, we will infer factors that inderlying mental illness. These factors are well constructed in transdiagnostic and RDoC framework.


\section{Social media study on mental health}


We introduce the 4P risks factors (predisposing, precipitating, perpectuating, protective) and the presenting factors in the previous sections. Now we group the social media studies on mental health based on these five components. 

\subsubsection{predisposing factors} (import citation list)

Social media studies that cover the predisposing factors include: gender and culture differences, psychological traits and self-disclosure cite(). The psychological traits study mainly include personality and self-esteem.

 \textbf{gender and culture differences.} Study shows that the dislosure of traumatic events various across cultures \cite{freitag2011talking,de2016quantifying}. De Choudhury et al., study the social media disclosure on mental illness and find significant differences between content shared by male and female users and users from different countries. Male users shows more negativity and more relunctant to seek social support. Mental health disclosure users from various cultures display different language patterns in psychological processing \cite{de_choudhury_gender_2017}. Park et al, found that female Facebook users are warmer, more compassionate and slightly more assertive in their language \cite{park_women_2016}.  Pennebaker et al, found that depressed individuals who wrote in Spanish showed more relational concern, whereas, individuals who wrote in English showed more concern in medication \cite{ramirez2008psychology}.


 \textbf{Personality.} Numourous studies have found relationship between user personality and social media usage \cite{rife_participant_2016,farnadi_computational_2016,ross_personality_2009,moore_influence_2012,back_facebook_2010,seidman_self-presentation_2013,hughes_tale_2012,hong_analysis_2014,skues_effects_2012,zhang_gratifications_2011,lee_personality_2014,jenkins-guarnieri_relationships_2012,blachnio_psychological_2013,wang2013share}. Hughes et al., found personality influence whether a person have a preference to use Facebook or Twitter \cite{hughes_tale_2012}. People with high conscientiousness tend to use Twitter, while sociable people tend to use Facebook \cite{ross_personality_2009}.  Extraverts are more active in Facebook activities while neuroticism is negatively correlated with the likelihood to writing comments. Chiu and colleagues found that self-inferiority predict Facebook use and people with a depressive character are prone to have Facebook addiction \cite{lee_personality_2014}. Chak and Leung also found that shyness predict Facebook addiction \cite{chak_shyness_2004}. Individuals high in openness are more likely to post and share intellectual topics \cite{marshall_big_2015}. People who are high in arcissism use Facebook for attention seeking. They are more likely to update their diet and exercise routine \cite{marshall_big_2015,andreassen_relationship_2017}.

 \textbf{self-presentation} Studies also find that self-presentation on Facebook is related to personality \cite{seidman_self-presentation_2013}. Lee and colleagues found that extroversion is positively related to self-presentation on Facebook \cite{lee_personality_2014}.  Seidman and colleagues found that low concientiousness and high neuroticism predicts self-presentation. High agreeableness and neuroticism predicts belongingness-related behaviours \cite{seidman_self-presentation_2013}.

Facebook profile reflects the actual self rather than the ideal self because the profile information predit observer impression rating but not the ideal-self rating. However, accuracy on people with high neuroticism was very low, which means that neuroticism is difficult to detect in stranger context \cite{funder1999personality,back_facebook_2010}.

\textbf{Self-esteem.}
Self-esteem is an important predisposing risk factor, high self-esteem has a protective effect on mental illness. Studies find that the frequency of using social media site is indirectly link to self-esteem and well-being. The frequency of using the site enhance the number of relationship formed on the site and the feedback received. Adolescents with positive feedback have an increased self-esteem, and vice versa \cite{valkenburg2006friend}.  Study also find that people with low self-esteem are more likely to update romantic partner \cite{marshall_big_2015}.


\textbf{self-disclosure.}
Self-disclosure is influential in the choice of social media platforms. Social media platforms such as Facebook and Twitter can be tie to a user's real-life social network when. Whereas, anonymous platforms such as reddit dissociate a person's real identity with the online community. Studies find that content intimacy not only regulate self-disclosure in face-to-face communication, but also in the social media settings \cite{ma_anonymity_2016}. Social media platforms that focus on anonymity enables users to have higher disclosure level \cite{andalibi_announcing_2018}. Studies find that factors that affect which social media platform to use lies in motivations. xx et al, found that motivation to communication predicts the preference on using Facebook while other motivations may be influential in the decision of which social media platforms to use \cite{jenkins2012relationships}


Hollenbaugh and colleagues studied how social media user behaviour reflect the amount, depth and breath of self-disclosure. They find that those who post a large amount of information about themselve utilize Facebook to maintain existing relationships. People who are lower in conscientiousness, agreeableness and social cohesion also tend to disclose more information. Those high in extroversion, less conscientious but more agreeable individuals provide greater depth in their disclosure. Whereas, those with lower self-esteem and less neuroticism disclose information that covers more varieties of topics \cite{hollenbaugh_facebook_2014}.


There are also studies that focus on how important or stressfull life events are disclosed on Facebook. It is found that people tend to share postive life events indirectly and negative life events directly \cite{bevan_how_2015}. In addition, age and loneliness also predict Facebook disclosure (Who discloses the most on Facebook \cite{malik2016uses}.

Whether one will disclose sensitive information on social media depends on the evalution of the feedback they will receive. Users weight the benefits and drawbacks before they are making sensitive disclosure on social media. Whether they will post sensitive disclosure depends on whether they perceive the discloure will mitigate or prolong the emotional pain. Gaining public support is the major motivtation to disclose sensitive information \cite{vitak2017benefits}. 


xx et al study the exposure level of information from different domains on Facebook and real life, and whether the information share on FB is consistent with real life. They found that the level of exposure on information about relationship issues, address, phone numbers, email address and health history are much lower than physical world. Whereas, photos about the user and the user's family much higher exposure on Facebook than in physical world. The exposure level of positive and negative emotion, places they visit, political views, relationship status, work and study places are similar on Facebook and physical world \cite{villela2015contrasting}.


\subsubsection{precipitating factors}


\textbf{Stress prediction models}. Stress is the major precipitating factor for mental illness. There is large amount of stuides that builds machine learning classifier to predict stress level based on social media data. The process of building stress prediction classifer mainly involve human labels on posts that shows stress then using positive and negative valence words as features in the machine learning models \cite{thelwall2017tensistrength,mogadala2012twitter,lin2014user}. 
Some of the studies tie the stress prediction classifer to an App that provides theraputic advices to the user in an attempt to improve their coping skills. (Helping Teenagers Relieve Psychological Pressures:A Micro-blog Based System). However, the theraputic effect of these apps are not fully studied yet.

\textbf{Life events}. 

Sentiment is a direct indicator of life stress, which indicate the perceived level of stress. Whereas stressful life events reflect stress from an objective perspective. Researchers are insterested in how different stressful life events are disclosed on social media. xx et al., suggest that people engage in breakup disclosure in two ways: one-to-many announcement on the wall of the social media platform or change private status, which reflects self-presentation \cite{haimson2018relationship}. Posts the share traumatic life events have a greater amount of negative emotions. There is cultural diversity in sharing traumatic events \cite{freitag2011talking}. Those who experience beareavement tend to use more first person singular pronouns, past tense verbs, adverbs, prepositions, conjunctions, and negations \cite{brubaker2012grief}.
 
 In order to study the impact of life events on people, first, reseachers need to identify life events from the text. Life events are difficult to identify from the text because users often report them indirectly, thus some of the life events need to be be inferred from the text. Bevan and colleagues found that positive events are usually shared indirectly but negative events are reported directly \cite{bevan2015important}. In addition, life events are very specific to the user, using a keyword search technique is not comprehensive. Nevertheless, many users do not report their life events on social platform. 

Despite the fact that life events are difficult to identify, there are a few studies build predition models on life events \cite{khodabakhsh2018detecting,abe2018predicting}. xx et al, (Detecting life events from twitter based on temporal semantic features) compare the sentence smilarity of daily life tweets with tweets that report life events and use their distances as features in the classifier. They also find that if a sequence of tweets occur before or after a tweet, the tweet is likely to be an important life event \cite{khodabakhsh2018detecting}. Dickinson and colleagues find that if users suddent comment on topics that they are not usually interested in, they are likely to experience a prominant life event related to the comment \cite{dickinson2015identifying}.

There are also studies that focus on stressful life events only. The dectection of stressful life events mainly use positive or negative emotion as features. Manual labels for distress content are required for,the training data. (cite). 


\subsubsection{perpectuating factors}

Perpetuating factors are risk factors that maintain the symptoms or behaviors. Stress can be chronic if the stress is bind with living environment, social roles or social status. As a result, perpetuating factors include chronic diseases, malnutrition, crowding, pollution, social stigma, social roles, culture, social status and so on. Some of the chronic diseases has been investigated using social media data (cite physical list). xx et al, study how diabetes patients seek help information on social media \cite{shaw2011health}.

Social stigma is one of the limitations in social media studies. Most of the people are unwilling to share health related information on social media due to stigma \cite{moorhead2013new}. Sometimes social roles and social status are revealed in the the user profile as demographic information. However, there is a limitation that the microlevel information about social role and social status often do not exist on social media platforms, because this information is highly relevant to self-presentation.


\subsubsection{protective factors}

protective factors include, social support, effective coping skills, high self-esteem, low sensitivity to stress and so on. A large amount of social media study on mental health focus on the social support provided and received on various social media platforms, such as Facebook and Twitter (Sharing feelings online: studying emotional well-being via automated text analysis of Facebook posts), Instagram (Sensitive Self-disclosures, Responses, and Social Support on Instagram: The Case of #Depression) and Reddit (cite reddit list). xx found that individuals with higher level of depression or anxiety level tend to post negative emotions on Facebook more frequently. Moreover, the use of emoticons with positive emotions reflect stress level. This phenomenon is more prominent among the younger age group.
(Sharing feelings online: studying emotional well-being via automated text analysis of Facebook posts). Study also find that people use Instagram to tell difficult experience. Personal narratives, food and beverage, references to illness, and self-appearance concerns receive more social support. However, there's little aggression or support to the harmful or pro-disease behaviours. (Motivations for sharing information and social support in social media: A comparative analysis of Facebook, Twitter, Delicious, YouTube, and Flickr)

Social media sites provide a platform for individuals to narrate their difficult experience. Narration and writing have been found to have theraputic effect for a long time cite(). This theraputic effect is also found in anonymous social media sites. Anonymous disclosure greatly enhance self-disclosure level. Study find that anonymous disclosures on other sites facilitate disclosure on sites that link to real life profile (e.g. Facebook) \cite{andalibi2018announcing}. Anonymous disclosure on social media has been found to have theraputic effect on individuals who experience crisis. XX el al compare individuals who comment on the counseling recommendations shows evidence of healing response to crisis compare with those who didn't receive counseling recommendations on social media.  (A Social Media Based Examination of the Effects of Counseling RecommendationsAfter Student Deaths on College Campuses) 

Other factors that influence the theraputic effect include: emotional competence and linguistic accomodation. Study found that emotional communication competence affect one's expression and reception of emotional support. Giving and receiving emotional support in supporting groups has positive affect on those high in emotional communication competence but detrimental to those with low emotional communication competence \cite{yoo2014giving}. XX et al., also find that the linguistics alignment with the online communities influence the social support received. Communities that cater to mental health problems or have sensitive topics develop their linguistics conventions, therefore, members in the group preceive that social support is available in the community. As a consequence, those who maintain linguistics alignment with the community receive more social support. However, a serious problem raised in here is that those with low linguistics alignment often show greater vulnerability and needs more help. As a result, future design and studies on social media should focus more on this subgroup. (Mental Health Support and its Relationship to Linguistic Accommodation in Online Communities)

\textbf{Types of social support.}
Multiple studies identified the type of social support mentioned in psychology literature on social media platforms. \cite{yoo2014giving}(Mental Health Support and its Relationship to
Linguistic Accommodation in Online Communities). XX and colleagues found that all communities on reddit offer both information and emotional support, but these communities cater to different social support goals. Some communities requested and provided more emotional support, such as Mood Disorder communities. Whereas, some communities requested and provided more informational support, such as Compulsive Disorder communities. (Mental Health Support and its Relationship to
Linguistic Accommodation in Online Communities) This finding matches with the social support literature in psychology, social support that matches with the needs has buffering effects to crisis.

----
(The Language of Social Support in Social Media and its Effect on Suicidal Ideation Risk)
Incorporating human assessments in a stratified propensity score analysis based framework, we identify comparable subpopulations of individuals and measure the effect of online social support language. We interpret these linguistic cues with an established theoretical model of social support, and find that esteem and network support play a more prominent role in reducing forthcoming risk. 

types of support and their tokens on reddit 
We observe that emotional support ranges from relating to specific challenging life situations (“totally agree”), to expressing empathy over how they impact one’s life (“not alone”). Esteem support can include boosting one’s morale to fight the stigma of mental illness (“fight the”), or encouraging hope and uplifting thoughts despite distressful experiences (“misery”). Through informational support, commenters provide advice and suggestions to seek professional mental health help (“be tough”), therapy and medication (“medication”). Commenters also provide various forms of instrumental support, including self-improvement activities like involvement in pastimes (“a hobby”). Next, network support comments express solidarity, connection, and social integration (“is helpful”). Finally, acknowledgments of social support include explicit recognition of the post author’s feelings (“be super”), thoughts, or experiences (“depressed”).

social network site as the surveilance tool
Study about the radar app:
(Looking out for each other online: Digital outreach, emotional surveillance and safe(r) spaces)
These include a need to be explicit about organisational assumptions about what constitutes support and how consent to be supported is sought and given online; being aware of the potential for outreach to be both silencing and amplifying and of the sensitivity of the tipping point between keeping a digital eye out and emotional surveillance; and recognising the specific risks of using algorithms in relation to the most human of activities - responding to emotional pain. While only implicitly present in this article, trust runs through all these considerations and is core to what constitutes feeling ‘safe(r). Drawing on the meaning of safe(r) spaces repositions the sharing of emotions as being about more than the commercial interests of social media platforms or the norms relating to Twitter, \textbf{and highlights that people share distress or emotions because of their sense of trust in relationships they believe they have or desire to have. Any organisation engaging with digital outreach needs to understand how their actions can either shore up or breach such trust, and hence vitally affect people's sense of being safe.}



(Major Life Changes and Behavioral Markers in Social
Media: Case of Childbirth)
for a certain number of mothers (14-
15\%), activity goes down, PA goes down, NA goes up,
activation and dominance go down together, and the use of
1st person pronouns goes up, while that of 3rd person
pronouns goes down. We also notice that certain mothers
consistently show these dynamics over the entire postnatal
period of our analysis. In essence, we find that there are
new mothers who are exhibiting signs of decreased social
interactions, as manifested through social media, along with
a number of changes in emotional expression in a generally
negative direction.



-------

social network, social ties(cite social network study, tie strength)

lonely  -> disclose more on FB -> bigger social network (less high quality social ties)


students with higher openness levels reported spending more time on Facebook and having more friends on Facebook. Interestingly, students with higher levels of loneliness reported having more Facebook friends. (Williams et al)


 Drawing on crisis and psychology research, we find that the exposed individuals show greater grief, psycholinguistic, and social expressiveness, providing evidence of a healing response to crisis and thereby positive psychological effects of the counseling recommendations. We discuss the implications of our work in supporting post-crisis rehabilitation and intervention efforts on college campuses. (Font Size: 
A Social Media Based Examination of the Effects of Counseling Recommendations after Student Deaths on College Campuses)




 Results indicate that the more time spent on and the more social network memberships, the higher stress and lower quality of life; Facebook-specific usage was unrelated to either well-being variable. Together, these findings suggest that the current increase in social media variety and usage may be detrimental to user well-being. Users who shared important, bad health news on Facebook had higher stress and lower quality of life than those who did not, with no significant differences for sharing good health news. The more that users did not share important news on Facebook for self-protection and friend unresponsiveness reasons, the greater their stress. The self-protection reason was also negatively related to quality of life. These inconsistent findings can likely be partially explained by the nature of the information that is shared. These findings are discussed in light of disclosure and relationship patterns on social networks. (Disclosures about important life events on Facebook: Relationships with stress and quality of life)

\section{presenting}

We find that increased social isolation and lowered availability of social capital on Facebook, are the best predictors of PPD in mothers. (Characterizing and Predicting Postpartum Depression from Shared Facebook Data)


(Major Life Changes and Behavioral Markers in Social
Media: Case of Childbirth)
for a certain number of mothers (14-
15\%), activity goes down, PA goes down, NA goes up,
activation and dominance go down together, and the use of
1st person pronouns goes up, while that of 3rd person
pronouns goes down. We also notice that certain mothers
consistently show these dynamics over the entire postnatal
period of our analysis. In essence, we find that there are
new mothers who are exhibiting signs of decreased social
interactions, as manifested through social media, along with
a number of changes in emotional expression in a generally
negative direction.

\section{Social media data analysis techniques}

low social status increases the risk of being exposed to a number of adverse conditions, both physical and psychological (Adler, Marmot, McEwen and Stewart, 1999; Adler and Matthews, 1994). 


-------
Conclusion of literature review 
A New Dimension of Health Care: Systematic Review of the Uses, Benefits, and Limitations of Social Media for Health Communication

( The six key overarching benefits were identified as (1) increased interactions with others, (2) more available, shared, and tailored information, (3) increased accessibility and widening access to health information, (4) peer/social/emotional support, (5) public health surveillance, and (6) potential to influence health policy. Twelve limitations were identified, primarily consisting of quality concerns and lack of reliability, confidentiality, and privacy.)


 the key recommendations for future health communication research focus on robust and comprehensive evaluation and review, using a range of methodologies. The research priorities are highlighted below:

To determine the impact of social media for health communication in specific population groups with large sample sizes (representation of population groups).
To determine the relative effectiveness of different social media applications for health communication using RCTs.
To determine the longer-term impact on the effectiveness of social media for health communication using longitudinal studies.
To explore potential mechanisms for monitoring and enhancing the quality and reliability of health communication using social media.
To investigate the risks arising from sharing information online and the consequences for confidentiality and privacy, coupled with developing the most suitable mechanisms to effectively educate users in the maintenance of their confidentiality and privacy.
To determine how social media can be effectively used to support the patient-health professional relationship.
To determine the impact of peer-to-peer support for the general public, patients, and health professionals to enhance their interpersonal communication.
To explore the potential for social media to lead to behavior change for healthy lifestyles to inform health communication practice.
(A New Dimension of Health Care: Systematic Review of the Uses, Benefits, and Limitations of Social Media for Health Communication)

----

what is missing? linking predisposing factors to mental health. What matters is stress interacting with predisposing factors and protective factors 
(personality, self-esteem (studies only find correlation of certain social media behaviors), culture, gender, protective: ***social ties, social support, social roles)




